---
output: html_document
editor_options: 
  chunk_output_type: console
---

Welcome to the hands-on demo of the `mmrm` package!

Let's load the package first.

```{r load-package}
library(mmrm)
```

# Simplest model fit

Here we fit a mmrm model with an unstructured (`us`) covariance structure:

```{r getting-started}
fit <- mmrm(
  formula = FEV1 ~ RACE + SEX + ARMCD * AVISIT + us(AVISIT | USUBJID),
  data = fev_data
)
```

From the print output we see the function call,
the data used, the covariance structure with number of variance parameters, as well
as the likelihood method, and model deviance achieved, estimated coefficients
and model convergence:

```{r print}
fit
```

The `summary()` method then provides the coefficients table with Satterthwaite
degrees of freedom as well as the covariance matrix estimate:

```{r summary}
summary(fit)
```

# Common customizations

## Control Function

For fine control, `mmrm_control()` can be used, for example:

```{r low-level-control}
mmrm_control(
  method = "Kenward-Roger",
  optimizer = c("L-BFGS-B", "BFGS"),
  n_cores = 2,
  start = c(0, 1, 1, 0, 1, 0),
  accept_singular = FALSE,
  drop_visit_levels = FALSE
)
```

Note that this control list can either be passed via the `control` argument
to `mmrm`, or selected controls can be directly specified in the `mmrm` call.
We will see this below.

## REML or ML

Let's use the non-default maximum likelihood (ML) estimation:

```{r common-changes-reml}
fit_ml <- fit |> update(reml = FALSE)
fit_ml
```

We see that we can use the `update()` method to quickly update from an existing fit.

## Optimizer

We can check which optimizers were used:

```{r check-initial-optimizer}
attributes(fit)
```

Sometimes you want to manually control the optimization algorithm (optimizer) that is used:

```{r common-changes-optim}
fit_opt <- fit |> update(optimizer = "BFGS")
```

And we can check that it was indeed used:

```{r check-manual-optimizer}
attributes(fit_opt)
```

## Covariance Structure

Using different abbreviations in the model formula we can change the assumed covariance structure.
For example here for homogeneous compound symmetry (`cs`):

```{r common-changes-cov}
fit_cs <- fit |> update(formula = FEV1 ~ RACE + ARMCD * AVISIT + cs(AVISIT | USUBJID))
fit_cs
```

For a full overview we can look at the help page:

```{r cov-overview}
?covariance_types
```

## Weighting

You can perform weighted inference by specifying a numeric vector `weights` with positive values:

```{r common-changes-weights}
fit_wt <- fit |> update(weights = fev_data$WEIGHT)
fit_wt
```

## Grouped Covariance Structure

Covariance matrices for each group are identically structured (unstructured, compound symmetry, etc) but the estimates are allowed to vary across groups.
We use the form `xy(time | group / subject)` to specify the group variable.

Here is an example of how we use `ARMCD` as group variable:

```{r group-cov}
fit_cs <- fit |> update(formula = FEV1 ~ RACE + ARMCD * AVISIT + cs(AVISIT | ARMCD / USUBJID))
```

We can see that the estimated covariance matrices are different in different `ARMCD` groups:

```{r inspect-group-cov}
VarCorr(fit_cs)
```

## Degrees of freedom adjustment method

We can change this with the `method` argument:

```{r kr}
fit_kr <- fit |> update(method = "Kenward-Roger")
```

Note that this requires `reml = TRUE`, i.e. Kenward-Roger adjustment is not
possible when using maximum likelihood inference.

Looking at the `summary` we see the method and the
correspondingly adjusted standard errors and degrees of freedom:

```{r kr_summary}
summary(fit_kr)
```

Other choices are listed in:

```{r method-help}
?mmrm_control
```

## Robust sandwich estimator 

When using a structured covariance model, the robust sandwich estimator ensures
that inference is unbiased even if the covariance model is misspecified.

We can do this with the `vcov` option:

```{r cs-emp}
fit_cs_emp <- fit_cs |> update(vcov = "Empirical")
```

And again we see this in the summary:

```{r cs-emp-results}
summary(fit_cs_emp)
```

# Ecosystem integration

## `broom`

`mmrm` contains S3 methods methods for `tidy`, `glance` and `augment` (that also work without loading `broom`), e.g.:

```{r tidy-fit}
fit |> tidy()
```

## `emmeans`

We use `emmeans` to calculate least square means:

```{r emmeans-example}
library(emmeans)
emmeans(fit, ~ ARMCD | AVISIT)
```

## `tidymodels`

```{r}
library(tidymodels)
```

### `parsnip`

We can specify the fit directly using the linear regression engine:

```{r parsnip-example}
linear_reg() |>
  set_engine("mmrm", method = "Satterthwaite") |>
  fit(FEV1 ~ RACE + ARMCD * AVISIT + us(AVISIT | USUBJID), fev_data)
```

### `workflows`

```{r workflows-example} 
mmrm_spec <- linear_reg() |>
  set_engine("mmrm", method = "Satterthwaite")

mmrm_wflow <- workflow() |>
  add_variables(outcomes = FEV1, predictors = c(RACE, ARMCD, AVISIT, USUBJID)) |>
  add_model(mmrm_spec, formula = FEV1 ~ RACE + ARMCD * AVISIT + us(AVISIT | USUBJID))

mmrm_wflow |>
  fit(data = fev_data)
```

### `recipes`

We can separate out the data preparation step from the modeling step using the `recipes` package. Here we are converting the `ARMCD` variable into a dummy variable and creating an interaction term with the new dummy variable and each visit.

```{r recipe-example}
mmrm_recipe <- recipe(FEV1 ~ ., data = fev_data) |>
  step_dummy(ARMCD) |>
  step_interact(terms = ~ starts_with("ARMCD"):AVISIT)
```

Using `prep()` and `juice()` we can see what the transformed data that will be used in the model fit looks like.

```{r juice-example}
mmrm_recipe |>
  prep() |>
  juice()
```

We can pass the covariance structure as well in the `set_engine()` definition. This allows for more flexibility on presetting different covariance structures in the pipeline while keeping the data preparation step independent.

```{r spec-with-cov}
mmrm_spec_with_cov <- linear_reg() |>
  set_engine(
    "mmrm",
    method = "Satterthwaite",
    covariance = as.cov_struct(~ us(AVISIT | USUBJID))
  )
```

We combine these steps into a workflow:

```{r combine-workflow}
(mmrm_wflow_nocov <- workflow() |>
  add_model(mmrm_spec_with_cov, formula = FEV1 ~ SEX) |>
  add_recipe(mmrm_recipe))
```

Last step is to fit the data with the workflow object

```{r fit-workflow}
(fit_tidy <- fit(mmrm_wflow_nocov, data = fev_data))
```

To retrieve the fit object from within the workflow object run the following

```{r hardhat-extract}
fit_tidy |>
  hardhat::extract_fit_engine()
```

## `insight`

```{r insight-examples}
library(insight)
get_df(fit)
get_varcov(fit)
```

## `parameters`

```{r parameters-example}
library(parameters)
parameters(fit)
se_satterthwaite(fit)
```

